---
title: 'Trustworthy or not? Classifying political news articles’ credibility with machine learning'
output: html_notebook
---

# Data preparation
```{r Load packages}
library(readr)
library(caret)
library(quanteda)
library(quanteda.textmodels)
library(quanteda.textplots)
library(quanteda.textstats)
library(lexicon)
```

```{r Read data}
data_fake <- read_csv('Fake.csv') |> mutate(type = 'fake')
data_real <- read_csv('True.csv') |> mutate(type = 'real') 
news <- rbind(data_real, data_fake) |> select(c(text, type))
```

```{r Randomize row order}
set.seed(666)
rows <- news |> nrow() |> sample()
news <- news[rows,]
```

```{r Build corpus}
corpus_all <- corpus(news, text_field = 'text')
```

```{r Preprocess text}
tokens_all <- corpus_all |>
 tokens(remove_punct = T, remove_numbers = T, remove_symbols = T, remove_url = T) |> #remove punctuations, numbers, symbols and URLs
  tokens_tolower() |> #convert to lowercase 
  tokens_remove(stopwords('en', 'nltk')) |> #remove stopwords in the list of Natural Language Toolkit (NLTK)
  tokens_remove(c('reuters', '@\\w+', "\'s"), valuetype = 'regex') |> 
  tokens_replace( #lemmatize
    pattern = hash_lemmas$token, 
    replacement = hash_lemmas$lemma) |>
  tokens_keep(min_nchar = 2) #keep words with at least two characters
``` 

```{r Subset tokens by news type}
tokens_real <- tokens_all |> 
  tokens_subset(type == 'real')
tokens_fake <- tokens_all |> 
  tokens_subset(type == 'fake')
```

# Corpus statistics
```{r Build DTM for credible news}
dfm_real <- tokens_real |> dfm()
```

```{r Build DTM for fake news}
dfm_fake <- tokens_fake |> dfm()
```

```{r Plot wordclouds}
set.seed(666)
textplot_wordcloud(dfm_real, max_words = 100)
textplot_wordcloud(dfm_fake, max_words = 100)
```

```{r Word frequency tables}
freqs_real <- textstat_frequency(dfm_real)
freqs_real |> head(20)
freqs_fake <- textstat_frequency(dfm_fake)
freqs_fake |> head(20)
```

```{r Create testing and training sets}
set.seed(999)
testset <- tokens_all |> docnames() |> sample(9000) #about 25% 
tokens_test <- tokens_all |> 
  tokens_subset(docnames(tokens_all) %in% testset)
tokens_train <- tokens_all |> 
  tokens_subset(!docnames(tokens_all) %in% testset)
factor_train <- docvars(tokens_train, 'type') |> as.factor()
factor_test <- docvars(tokens_test, 'type') |> as.factor()
```

```{r Frequency tables by news type}
table(factor_train)
table(factor_test)
```

```{r Build DTM training set}
dfm_train <- tokens_train |> dfm()
```

```{r Build DTM testing set}
dfm_test <- tokens_test |> dfm() |>
  dfm_match(featnames(dfm_train))
```

# Machine learning
```{r Train Naive Bayes model}
train_nb <- textmodel_nb(dfm_train, factor_train)
summary(train_nb)
```

```{r Test Naive Bayes Model}
predict_nb <- predict(train_nb, newdata = dfm_test)
```

```{r Compute predictive accuracy}
mean(predict_nb == factor_test)
```

```{r Show confusion matrix and statistics}
confusion_matrix <- confusionMatrix(predict_nb, factor_test, mode = 'everything')
confusion_matrix
```

Get a list of Probability of Class Given Word scores (script by Wouter van Atteveldt!)
```{r PCGW score}
pc <- train_nb$priors |> as_tibble(rownames = 'class') |> 
  rename(pc=value)
# Get the p(w|c) from the params
pcgw <- t(train_nb$param) |> 
  as_tibble(rownames = 'word')  |>
  # Pivot to wide and add p(c)
  pivot_longer(-word, names_to = 'class', values_to = 'pwgc')  |>
  inner_join(pc, by = character()) |> 
  # Compute p(w)=sum(p(w|c)*p(c) and p(c|w)=p(w|c)*p(c)/p(w):
  group_by(word) |>
  mutate(pw = sum(pwgc*pc), 
         pcgw = pwgc*pc/pw)
pcgw <- pcgw |> arrange(-pcgw)
```

Create list of prediction errors & corpus containing only errors
```{r Prediction errors}
errors <- which(predict_nb != factor_test)
error_names <- docnames(dfm_test)[errors]
error_corpus <- corpus_all[error_names]
```

Get false positives and false negatives
```{r Type I and II errors}
false_fake <- corpus_subset(error_corpus, type == 'real')
false_true <- corpus_subset(error_corpus, type == 'fake')
```

Create data frame of the test set containing doc ID, article text, type, prediction and variable marking if prediction was correct
```{r}
news_test <- corpus_all |>
  corpus_subset(docnames(corpus_all) %in% testset)
predictions <- tibble(
  doc_id = names(predict_nb), 
  prediction = predict_nb)
actual_test <- tibble(
  doc_id = names(news_test), 
  type = factor_test)
error_data <- merge(actual_test, predictions)
```

```{r}
texts <- tibble(
  doc_id = docnames(news_test), 
  text = as.character(news_test)
  )
error_texts <- right_join(texts, error_data) |> 
  mutate(
    error = ifelse(
      type == prediction, 
      'correct', 
      'wrong')
    )
```


